\Sof{TBA - New Titles in the comments}

### 2020-05-04 paper/icsme20.tex
new update

The BCH tool does not compute the final score that our study needs to compare
maintainability amongst different project versions. For this, we
follow previous work on measuring the impact of energy-oriented refactorings
~\cite{cruz2019energyoriented}. Luis et al. ($2019$) proposed an equation
to capture the distance between the current state of the project and
the standard thresholds calculated by the BCH based on the insights 
provided in~\cite{Olivari:2018}. The equation provided in~\cite{cruz2019energyoriented} 
considers that the size of project changes does not affect the maintainability and
that the distance to lower severity levels is less penalized than to the
thresholds in high severity levels.
% 	thresholds in high severity levels
% 	difference between two versions
% The equation considers the following:
% \begin{itemize}
% 	\item \textbf{Size of project changes does not affect the maintainability
% 	difference between two versions, $\Delta M (v_{s-1},v_{s}) = M(v_{s}) - M(v_{s-1})$.} We
% 	aim at evaluating security patterns occurring in different projects similarly.
% 	Thus, the derived metric uses the \textit{raw} number of lines of code rather
%   percentages for normalization purposes.
% 	\item \textbf{Distance to lower severity levels is less penalized than to the
% 	thresholds in high severity levels.} Severity level weights based on the
% 	severity level to count lines of code that violate maintainability guidelines.
% \end{itemize}

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
\Sof{The compliance for a severity level
is calculated based on the number of lines of code that comply and not comply
with the guideline at a given severity level as explained in more
detail in previous work~\cite{cruz2019energyoriented}.} In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability. 
% The compliance $C$ for a given severity
% level $l$ is derived by:
%
% \begin{equation}\label{eq:3}
%     C(l) = LOC_{compliant}(l) - w(l) * LOC_{\neg compliant}(l)
% \end{equation}
%
% \noindent
% where $LOC_{compliant}(l)$ are the lines of code that comply with the guideline
% at the given severity level $l$, $LOC_{\neg compliant}(l)$ are the lines of code
% that do not comply with the guideline at the given severity level $l$ and $w(l)$
% is the weight factor to heighten the impact of non-compliant lines in comparison to
% compliant lines. Finally, the term $w(l)$ is calculated as follows:
%
% \begin{equation}
%     w(l) = \frac{1 - \theta(l)}{\theta(l)}
% \end{equation}
%
% \noindent
% where $\theta(l)$ is the threshold in percentage of the lines of code that are
% accepted to be non-compliant with the guideline for the severity level $l$. This
% is a standard value defined by BCH. In other words, the factor $w$ is used in
% Equation~\ref{eq:3} to highlight the lines of code that are not complying with
% the guideline. Then, we compute the difference of maintainability between the
% security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in
% Figure~\ref{fig:commit}.

@@

This study evaluates a final total of $998$ security and baseline commits from $262$ 
distinct open-source projects. The following section presents the results 
obtained for each research question.

\textit{\textbf{RQ1: What is the impact of security patches on the
maintainability of open-source software?}}

Maintainability is being calculated as the mean of the guidelines computed
by BCH. Figure~\ref{fig:guidelines} shows the impact of security patches on 
each guideline individually. For each type of guideline, three types of 
impact of patches in the software maintainability are presented: 
\emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is presented the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.

Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $6$ guidelines: 
\emph{Write Clean Code} (p-value = $3.66$x$10^{-51}$), \emph{Keep 
Architecture Components Balanced} (p-value = $5.07$x$10^{-73}$), 
\emph{Couple Architecture Components Loosely} (p-value = $1.61$x$10^{-08}$), 
\emph{Separate Concerns in Modules} (p-value = $0.090$), \emph{Keep Unit 
Interfaces Small} (p-value = $8.76$x$10^{-15}$), \emph{Write Code Once} 
(p-value = $4.97$x$10^{-35}$). Overall, for each of the $8$ guidelines we also 
observe a substantial number of cases where the maintainability suffers a 
negative impact $-$ from $10\%$ to $40\%$ $-$ which highlights the need of 
better practices/techniques/tools to produce software. However, the results do 
not provide statistical significance to the \emph{Separate 
Concerns in Modules} guideline - results should be read carefully.

According to the results, the \emph{Write Simple Units of Code} 
(p-value=$0.037$ and $372$ ($37.3\%$) of negative cases) and \emph{Write Short 
Units of Code} (p-value=$0.028$ and $375$ ($37.6\%$) of negative cases) are 
the most affected guidelines.

 For projects with large 
 codebases, the results retrieved for \emph{Keep Your Codebase Small} were not 
 viable because they retrieved values above the limit set by BCH ($20$ Person-years).
 The \emph{Automated Tests} guideline was also not considered because the BCH does not return
 the amount of lines of code that complies with the guideline -- which is one of the variables
 our formula uses to calculate the maintainability of the project regarding each guideline. 
 Thus, we decided to not consider both guidelines in our analysis.

 \begin{figure}[h]
  	\centering
 	\vspace{-0.3 cm}
  	\includegraphics[width=0.35\textwidth]{figures/main_per_guideline.pdf}
 	\vspace{-0.3 cm}
  	\caption{Impact of the security patches per guideline.}
 	\label{fig:guidelines}
 	\vspace{0.2 cm}
 	\scalebox{0.5}{
 	\begin{tabular}{r@{: }l r@{: }l}
 	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
 	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns in Modules \\
 	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
 	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of Code
 	\end{tabular}
 	}	
 \end{figure}

Some of the vulnerabilities are identified with \emph{Common Vulnerabilities 
and Exposure} (CVE) entries. We leveraged the \emph{National Vulnerability 
Database} (NVD) website to obtain their severity levels. In total, we 
retrieved severity scores for $551$ vulnerabilities: $117$ (\emph{HIGH}), 
$404$ (\emph{MEDIUM}) and $30$ (\emph{LOW}). Figure~\ref{fig:severity} 
presents the impact of security patches per severity level on the 
maintainability of open-source software. We observe that patches for 
\emph{HIGH} ($47.9\%$) and \emph{MEDIUM} ($44.1\%$) severity vulnerabilities 
hinder more the maintainability of software than \emph{LOW} ($30.0\%$) 
severity vulnerabilities. Again, patches have a considerable 
negative impact on software maintainability - between $30\%$ 
and $50\%$. None statistical significance was retrieved. However, 
results should not be disregarded because they seem to confirm the 
assumption that higher severity vulnerabilities patches may have a 
more negative impact on maintainability.

@@

We have also evaluated the impact on the maintainability per programming 
language of open-source software (Figure~\ref{fig:lang_main}). We restrict 
this analysis to programming languages with at least $20$ data points, as this 
is a requirement for our hypothesis tests. Thus, we compare the results for \texttt{C/C++}, 
\texttt{Ruby}, \texttt{Java}, \texttt{Objective C/C++}, \texttt{Python} and \texttt{PHP}. 
The remaining languages are included into \texttt{Other}.

\texttt{C/C++} ($46.54\%$), \texttt{Ruby} ($46.48\%$) and \texttt{PHP} ($38.27\%$) are
the programming languages with a more negative impact on the maintainability 
of software. \texttt{Java} seems to be the language less affected
by patching, more than $50\%$ of patches had a positive impact on maintainbility. 
But overall patches in all languages have a considerable
negative impact on maintainability - from $30\%$ to $50\%$ - which confirms
the need for better and more secure programming languages. 
Statistical significance was only retrieved for the \texttt{C/C++} 
(p-value = $2.24$x$10^{-05}$) and \texttt{Ruby} (p-value = $0.041$) languages.
Results should be read with care where statistical significance was not found.
Yet, data reports very interesting hints on the impact of programming languages on 
security patches.

@@

Regarding security commits results (Figure~\ref{fig:secvsreg}), we observe that
maintainability decreases in $41.28\%$ ($412$), remains the same in $20.64\%$ ($206$)
and increases in $38.08\%$ ($380$) of the cases. The resulting
p-value of the Paired Wilcoxon signed-rank test is $0.044$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact on the maintainability of
open-source software}.
%

@@

Results show that regular changes typically have no impact on the 
software maintainability which contradicts previous research.
However, we manually inspected a considerable amount of cases from
the regular changes distribution where this happens and found that 
identifying regular commits as similar in size as the security-related 
commit is limiting the type of regular commits (e.g, input 
refactorings, variables or functions, type conversion, etc).  
One example is the regular commit\footnote{Regular change: \url{https://github.com/AFNetworking/AFNetworking/compare/9cde4e45845486bdb46ad322448f93df3b8479bf...3486a008a1fd1348d0f2fe748a5e23933d94647a}} generated 
by our methodology and using as reference the following security commit
\footnote{Security change: \url{https://github.com/AFNetworking/AFNetworking/compare/49186d9a3032dc02da1d0cc445cbf69b0cfb001b...2b2558a90e7bfb2ef5169b1ca21b91851e5a5b7d}}. From a
overflow vulnerability patch, our methodology chose a similiar
type of change. We strongly believe that this phenomenon lead
that increased significantly the number of cases with no impact
on the software maintainability.

The resulting p-value of the Paired Wilcoxon signed-rank test 
is $0.331$, i.e., conclusions regarding the regular changes 
are not statistically significant. Yet, it is important to
note that \emph{patching security vulnerabilities seem to have a 
more negative impact on the maintainability of open-source 
software than performing regular changes}.

@@

\section{Discussion}\label{sec:discussion}

In this section, we discuss the results and answer the proposed research questions.

The guideline analysis suggests that all guidelines are affected by security patches. Being the
\emph{Write Short Units of Code}, \emph{Write Simple Units of Code} and \emph{Separate Concerns in Modules} the guidelines with higher negative impact on the maintainability ($>30\%$ of cases). \emph{Keep Architecture Components Balanced} is the guideline with lower negative impact on 
the maintainability (approx. $10\%$). This might imply that developers should give different
levels of attention to the guidelines when patching vulnerabilities.
%
As expected, vulnerabilities with higher severity are more likely to impact negatively
the maintainability of software. This suggests that high/medium severity vulnerabilities may
need more attention than low severity when patching vulnerabilities.
%

We would expect that the negative impact of programming languages on
maintainability to be more severe, as arguably poor design of programming
languages and lack of using the best practices lead to more buggy and vulnerable
code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. However,
Figure~\ref{fig:lang_main} shows that only \emph{C/C++} and \emph{Python} have a negative impact approximate to $50\%$ on software maintainability. We suspect that these values are
the result of projects contributions policies. Nine of the top-10 of dataset projects 
with more contributors have restrict contribution policies regarding code
standards. Some examples are the \texttt{laravel/framework},
\texttt{php/php-src}, \texttt{openssl/openssl} and \texttt{cakephp/cakephp}. For
instance, \texttt{laravel/framework} requires the \texttt{PSR-2
Standard}\footnote{PSR-2 Standard available at
\url{https://www.php-fig.org/psr/psr-2/} (Accessed on \today{})} that is a coding style guide for \texttt{PHP}.

\subsubsection*{\textit{\textbf{RQ2} \textbf{Which patterns of security patches are more likely to
affect open-source software maintainability?}}}

\textit{Path Traversal (CWE-22), Information Exposure (CWE-200), Resource Management Errors (CWE-399) are more likely to affect
open-source software maintainability.}

Overall, the results obtained for the maintainability per CWE are no statistical significance 
The impact of a patch depends on its complexity, i.e., if the patchs
adds complexity to the code base it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting} (CWE-79)
patches endure more
cases without any impact in the maintainability. Typically, these types of
vulnerabilities do not need extra lines to be fixed, as shown in
Listing~\ref{lst:fix}. Whereas, \emph{Resource Management Errors} (CWE-399) and vulnerabilities
such as the one presented in Section~\ref{sec:motivation} responsible by Denial-of-Service
attacks are more difficult to fix (several code lines deleted and added).
Another example is the Apple authentication flaw discovered in $2017$
(CVE-$2017$-$13872$)\footnote{CVE-$2017$-$13872$ details available at
\url{https://support.apple.com/en-us/HT208315} (Accessed on \today{})}, where any user
could log in as root with an empty password. Patrick Wardle examined the cause
of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at
\url{https://objective-see.com/blog/blog\_0x24.html} (Accessed on \today{})} and
concluded that the flaw was due to the introduction of high cyclomatic complexity
in a method used to verify the password. In sum, \emph{Cross-Site Scripting} (CWE-79) should take part of the
developers' preoccupations along with the other patterns that explicitly hinder
maintainability because despite their low complexity these are patterns that
still produce a considerable negative impact on the maintainability. Real
efforts have been made in creating and incorporating coding standards and best
practices in software production. However, security is still far from perfection
and fixing vulnerabilities might hinder software maintainability. Thus, it is
not only important to provide tool support to developers to help them apply these
patterns without harming software maintainability, but also to create and redesign
programming languages to facilitate developers improving software maintainability.

We measured the effect sizes between both distribution -- before and after
patches --- per language and security patterns using the Common Language Effect
Size~\cite{cliff:1993} method. Effect sizes are of a low-medium level.
Thus, considering the $0$ median and the low effect sizes, it is reasonable
to affirm that there is a significant effect but it might only be
recognized after a thoughtful data analysis.

\subsubsection*{\textit{\textbf{RQ3} \textbf{What is the impact of security patches versus regular changes
on the maintainability of open-source software?}}}
\textit{Patching security vulnerabilities has a larger negative and positive impact
than when fixing regular commits.}

We found that $41.76\%$ of the security patches have a negative impact on the
maintainability of open-source software (cf. Figure~\ref{fig:secvsreg}). While $21.04\%$ of regular commits have a negative impact on the maintainability. This raises
a new trade-off when developers need to patch vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Documentation} of software must provide developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use.

	\item\textbf{Programming Languages} should provide new design patterns to
	easily implement security patterns without endangering maintainability.
	Ultimately, new programming languages, both secure and maintainable by design,
	may be designed to help developers be highly productive when writing and
	maintaining  secure applications. One example is the need for designing new
	authorization mechanisms since these are one of the most complex security
	features to implement.

>>>>

Patches can be performed through one commit, several consecutive commits, or 
commit(s) interleaved with more programming activities, being the later
commonly called a floss-refactoring. This study considers mainly 1-commit 
patches. Yet, a small percentage of data points had more than one commit 
involved in the patch. We manually inspected these cases and $23$ 
floss-refactorings were identified and disregarded since it would not
be fair to measure the maintainability of these cases where other
programming activities are involved.

The BCH tool does not compute the final score that our study needs to compare
maintainability amongst different project versions. For this, we follow previous 
work on measuring the impact of energy-oriented 
refactorings~\cite{cruz2019energyoriented}. Luis et al. ($2019$) proposed an 
equation to capture the distance between the current state of the project and
the standard thresholds calculated by the BCH based on the insights provided 
in~\cite{Olivari:2018}. The equation provided in~\cite{cruz2019energyoriented} 
considers that the size of project changes does not affect the maintainability 
and that the distance to lower severity levels is less penalized than to the
thresholds in high severity levels.

% 	thresholds in high severity levels
% 	difference between two versions
% The equation considers the following:
% \begin{itemize}
% 	\item \textbf{Size of project changes does not affect the maintainability
% 	difference between two versions, $\Delta M (v_{s-1},v_{s}) = M(v_{s}) - M(v_{s-1})$.} We
% 	aim at evaluating security patterns occurring in different projects similarly.
% 	Thus, the derived metric uses the \textit{raw} number of lines of code rather
%   percentages for normalization purposes.
% 	\item \textbf{Distance to lower severity levels is less penalized than to the
% 	thresholds in high severity levels.} Severity level weights based on the
% 	severity level to count lines of code that violate maintainability guidelines.
% \end{itemize}

@@

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
\Sof{The compliance for a severity level
is calculated based on the number of lines of code that comply and not comply
with the guideline at a given severity level as explained in more
detail in previous work~\cite{cruz2019energyoriented}.} In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability.

\Sof{@todo: Don't forget to mention the limitations of BCH and the outliers.}

% The compliance $C$ for a given severity
% level $l$ is derived by:
%
% \begin{equation}\label{eq:3}
%     C(l) = LOC_{compliant}(l) - w(l) * LOC_{\neg compliant}(l)
% \end{equation}
%
% \noindent
% where $LOC_{compliant}(l)$ are the lines of code that comply with the guideline
% at the given severity level $l$, $LOC_{\neg compliant}(l)$ are the lines of code
% that do not comply with the guideline at the given severity level $l$ and $w(l)$
% is the weight factor to heighten the impact of non-compliant lines in comparison to
% compliant lines. Finally, the term $w(l)$ is calculated as follows:
%
% \begin{equation}
%     w(l) = \frac{1 - \theta(l)}{\theta(l)}
% \end{equation}
%
% \noindent
% where $\theta(l)$ is the threshold in percentage of the lines of code that are
% accepted to be non-compliant with the guideline for the severity level $l$. This
% is a standard value defined by BCH. In other words, the factor $w$ is used in
% Equation~\ref{eq:3} to highlight the lines of code that are not complying with
% the guideline. Then, we compute the difference of maintainability between the
% security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in
% Figure~\ref{fig:commit}.

@@

This study evaluates a final total of $998$ security and baseline commits 
from $262$ distinct open-source projects. The following section presents and 
discusses the results for each question. Research questions are also answered.

In this section, we report and discuss the impact of patches on open-source
software maintainability under four different views: guidelines/metrics, 
severity, programming language and formula proposed - the mean of all 
guidelines/metrics.

% As expected, vulnerabilities with higher severity are more likely to impact negatively
% the maintainability of software. This suggests that high/medium severity vulnerabilities may
% need more attention than low severity when patching vulnerabilities.
% %
% We would expect that the negative impact of programming languages on
% maintainability to be more severe, as arguably poor design of programming
% languages and lack of using the best practices lead to more buggy and vulnerable
% code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. However,
% Figure~\ref{fig:lang_main} shows that only \emph{C/C++} and \emph{Python} have a negative impact approximate to $50\%$ on software maintainability. We suspect that these values are
% the result of projects contributions policies. Nine of the top-10 of dataset projects
% with more contributors have restrict contribution policies regarding code
% standards. Some examples are the \texttt{laravel/framework},
% \texttt{php/php-src}, \texttt{openssl/openssl} and \texttt{cakephp/cakephp}. For
% instance, \texttt{laravel/framework} requires the \texttt{PSR-2
% Standard}\footnote{PSR-2 Standard available at
% \url{https://www.php-fig.org/psr/psr-2/} (Accessed on \today{})} that is a coding style guide for \texttt{PHP}.

\textbf{Guidelines/Metrics:} Each patch performs a set of changes
on the source code of the software. These changes may have a different
impact on the guidelines/metrics used to measure software maintainability. 
Figure~\ref{fig:guidelines} shows the impact of security patches 
on each guideline individually. Under each guideline is stated the 
metric used to the calculations. Table~\ref{tab:guidelines} describes 
in more detail the metrics behind the guidelines presented. 
For each type of guideline, three types of impact of patches in 
the software maintainability are presented: 
\emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is presented the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.

Regarding the impact of security patches for each guideline, we observe
that overall patches have a more positive impact than negative in
all guidelines/metrics. However, patching vulnerabilities also has a very
significant negative impact on all guidelines - between $10\%$ and $40\%$. 
\emph{Write Simple Units of Code} ($37.3\%$), \emph{Write Short Units of 
Code} ($37.6\%$) and \emph{Separate Concerns in Modules} ($33.2\%$) are the 
most affected guidelines. This may imply that developers when patching 
vulnerabilities fail to design/implement solutions that continue to respect 
the limit bounds of branch points (/programming conditions) and function/module 
sizes (e.g., functions should not have more than $15$ LOCs). Still, on 
respecting bound limits, developers also seem to not consider the limit 
of $4$ parameters per function for the \emph{Keep Unit Interfaces Small} 
guideline required by BCH, in $24.5\%$ of the cases. This guideline is usually
violated when the patch requires to input new information to a function/class 
and developers fail to use the \emph{Introduce Parameter Object} refactoring 
pattern. Results do not provide statistical significance to the \emph{Separate 
Concerns in Modules} guideline, i.e., results should be read carefully.

Software architecture is also affected while patching vulnerabilities.
Both \emph{Couple Architecture Component Loosely} and \emph{Keep
Architecture Components Balanced} guidelines suffer a negative impact of 
$26.7\%$ and $11.1\%$, respectively. Component independence and balance
are important to make easier to find the source code developers
want to patch/improve and to understand how the high-level components
interact with others. However, results may imply that developers
fail to use techniques such as encapsulation to hide implementation
details and make the system more modular.

Software systems typically have $9\%$-$17\%$ of cloned code~\cite{5773403}. 
The \emph{Write Code Once} guideline results show that duplicated code 
increased in $174$ ($17.4\%$) of the patches. Previous work showed a 
correlation between code smells and code duplication~\cite{7476787} 
which may also be reflected in the \emph{Write Clean Code} guideline results. 
BCH reported new code smells for $152$ ($15.2\%$) patches which according 
to previous work may harm software security since some of these new code
smells may be new vulnerabilities~\cite{8819456}. Thus, while developers 
tend to ignore refactoring techinques such as the \emph{Replace Method 
with Method Object} technique, they may also be introducing new software 
vulnerabilities.

For projects with large codebases, the results retrieved for \emph{Keep Your 
Codebase Small} were not viable because they retrieved values above the limit 
set by BCH ($20$ Person-years). The \emph{Automated Tests} guideline was also 
not considered because the BCH tool does not return the same type of information
as for other guidelines. Due to BCH limitations, we detected a few data points
that retrieved incorrect overall calculations for one version of the software. 
Those data points were disregarded from the study.

Although overall patching vulnerabilities has a less negative impact, it is 
important to note that the design debt of one guideline can lead
to severe impacts on the software quality~\cite{10.1145/1985362.1985366} 
and in the market value of a company~\cite{4267025}. Thus,
tools like Better Code Hub and static analysis should be leveraged to 
assess the risk of patching vulnerabilities - alongside with pull
requests and code reviews. While style-guided static code analysis
may be daunting due to the amount of rules and different effects
on maintainability, BCH ensures that the $10$ chosen guidelines are
the ones with the highest effect on maintainability. 
%
% 
 \begin{figure}[H]
  	\centering
 	\vspace{-0.3 cm}
  	\includegraphics[width=0.45\textwidth]{figures/main_per_guideline.pdf}
 	\vspace{-0.3 cm}
  	\caption{Impact of the security patches per guideline.}
 	\label{fig:guidelines}
 	\vspace{0.2 cm}
 	\scalebox{0.5}{
 	\begin{tabular}{r@{: }l r@{: }l}
 	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
 	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns 
	in Modules \\
 	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
 	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of 
	Code
 	\end{tabular}
 	}	
 \end{figure}

\textbf{Severity:} Some of the vulnerabilities are identified with \emph{Common 
Vulnerabilities and Exposure} (CVE) entries. We leveraged the \emph{National Vulnerability 
Database} (NVD) website to obtain their severity levels. In total, we 
retrieved severity scores for $551$ vulnerabilities: $117$ (\emph{HIGH}), 
$404$ (\emph{MEDIUM}) and $30$ (\emph{LOW}). Figure~\ref{fig:severity} 
presents the impact of security patches per severity level on the 
maintainability of open-source software. We observe that patches for 
\emph{HIGH} ($47.9\%$) and \emph{MEDIUM} ($44.1\%$) severity vulnerabilities 
hinder more the maintainability of software than \emph{LOW} ($30.0\%$) 
severity vulnerabilities. Again, patches have a considerable 
negative impact on software maintainability - between $30\%$ 
and $50\%$. None statistical significance was retrieved. However, 
results should not be disregarded because they seem to confirm the 
assumption that higher severity vulnerabilities patches may have a 
more negative impact on maintainability.

@@

\textbf{Programming Languages:} We have also evaluated the impact on the maintainability per programming 
language of open-source software (Figure~\ref{fig:lang_main}). We restrict 
this analysis to programming languages with at least $20$ data points, as this 
is a requirement for our hypothesis tests. Thus, we compare the results for 
\texttt{C/C++}, \texttt{Ruby}, \texttt{Java}, \texttt{Objective C/C++}, 
\texttt{Python} and \texttt{PHP}. The remaining languages are included into 
\texttt{Other}. \texttt{C/C++} ($46.54\%$), \texttt{Ruby} ($46.48\%$) and 
\texttt{PHP} ($38.27\%$) are the programming languages with a more negative
impact on the maintainability of software. \texttt{Java} seems to be the 
language less affected by patching, more than $50\%$ of patches had a positive 
impact on maintainbility. But overall patches in all languages have a 
considerable negative impact on maintainability - from $30\%$ to $50\%$ - which 
confirms the need for better and more secure programming languages. 
Statistical significance was only retrieved for the \texttt{C/C++} 
(p-value = $2.24$x$10^{-05}$) and \texttt{Ruby} (p-value = $0.041$) languages.
Results should be read with care where statistical significance was not found.
Yet, data reports very interesting hints on the impact of programming languages 
on security patches.

@@

\textbf{Formula:} Maintainability is being calculated as the mean of 
the guidelines computed by Better Code Hub (BCH). Regarding security 
commits results (Figure~\ref{fig:secvsreg}), we observe that
maintainability decreases in $41.28\%$ ($412$), remains the same in $20.64\%$ 
($206$) and increases in $38.08\%$ ($380$) of the cases. The resulting
p-value of the Paired Wilcoxon signed-rank test is $0.044$. Since the p-value is 
below the significance level of $0.05$, we argue that \emph{security patches 
have a negative impact on the maintainability of open-source software}.
%

@@

Results show that regular changes typically have no impact on the 
software maintainability which contradicts previous research.
However, we manually inspected a considerable amount of cases from
the regular changes distribution where this happens and found that 
identifying regular commits as similar in size as the security-related 
commit is limiting the type of regular commits (e.g, input 
refactorings, variables or functions, type conversion, etc).  
One example is the regular commit\footnote{Regular change: 
\url{https://github.com/AFNetworking/AFNetworking/compare/9cde4e45845486bdb46ad32
2448f93df3b8479bf...3486a008a1fd1348d0f2fe748a5e23933d94647a}} generated 
by our methodology and using as reference the following security commit
\footnote{Security change: 
\url{https://github.com/AFNetworking/AFNetworking/compare/49186d9a3032dc02da1d0cc
445cbf69b0cfb001b...2b2558a90e7bfb2ef5169b1ca21b91851e5a5b7d}}. From a
overflow vulnerability patch, our methodology chose a similiar
type of change. We strongly believe that this phenomenon lead
that increased significantly the number of cases with no impact
on the software maintainability.

The resulting p-value of the Paired Wilcoxon signed-rank test is $0.331$, i.e., 
conclusions regarding the regular changes are not statistically significant. 
Yet, it is important to note that \emph{patching security vulnerabilities seem 
to have a more negative impact on the maintainability of open-source software 
sthan performing regular changes}.

@@

% \section{Discussion}\label{sec:discussion}
% In this section, we discuss the results and answer the proposed research
% questions.
% \textit{\textbf{RQ1: What is the impact of security patches on the
% maintainability of open-source software?}}
% \textit{Patching vulnerabilities may harm the maintainability of open-source
% software.}
% The guideline analysis suggests that all guidelines are affected by security patches. Being the
% \emph{Write Short Units of Code}, \emph{Write Simple Units of Code} and \emph{Separate Concerns in Modules} the guidelines with higher negative impact on the maintainability ($>30\%$ of cases). \emph{Keep Architecture Components Balanced} is the guideline with lower negative impact on
% the maintainability (approx. $10\%$). This might imply that developers should give different
% levels of attention to the guidelines when patching vulnerabilities.
% %
% As expected, vulnerabilities with higher severity are more likely to impact negatively
% the maintainability of software. This suggests that high/medium severity vulnerabilities may
% need more attention than low severity when patching vulnerabilities.
% %
% We would expect that the negative impact of programming languages on
% maintainability to be more severe, as arguably poor design of programming
% languages and lack of using the best practices lead to more buggy and vulnerable
% code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. However,
% Figure~\ref{fig:lang_main} shows that only \emph{C/C++} and \emph{Python} have a negative impact approximate to $50\%$ on software maintainability. We suspect that these values are
% the result of projects contributions policies. Nine of the top-10 of dataset projects
% with more contributors have restrict contribution policies regarding code
% standards. Some examples are the \texttt{laravel/framework},
% \texttt{php/php-src}, \texttt{openssl/openssl} and \texttt{cakephp/cakephp}. For
% instance, \texttt{laravel/framework} requires the \texttt{PSR-2
% Standard}\footnote{PSR-2 Standard available at
% \url{https://www.php-fig.org/psr/psr-2/} (Accessed on \today{})} that is a coding style guide for \texttt{PHP}.
% \subsubsection*{\textit{\textbf{RQ2} \textbf{Which patterns of security patches are more likely to
% affect open-source software maintainability?}}}
% \textit{Path Traversal (CWE-22), Information Exposure (CWE-200), Resource Management Errors (CWE-399) are more likely to affect
% open-source software maintainability.}
% Overall, the results obtained for the maintainability per CWE are no statistical significance
% The impact of a patch depends on its complexity, i.e., if the patchs
% adds complexity to the code base it is probably affecting the software
% maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting} (CWE-79)
% patches endure more
% cases without any impact in the maintainability. Typically, these types of
% vulnerabilities do not need extra lines to be fixed, as shown in
% Listing~\ref{lst:fix}. Whereas, \emph{Resource Management Errors} (CWE-399) and vulnerabilities
% such as the one presented in Section~\ref{sec:motivation} responsible by Denial-of-Service
% attacks are more difficult to fix (several code lines deleted and added).
% Another example is the Apple authentication flaw discovered in $2017$
% (CVE-$2017$-$13872$)\footnote{CVE-$2017$-$13872$ details available at
% \url{https://support.apple.com/en-us/HT208315} (Accessed on \today{})}, where any user
% could log in as root with an empty password. Patrick Wardle examined the cause
% of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at
% \url{https://objective-see.com/blog/blog\_0x24.html} (Accessed on \today{})} and
% concluded that the flaw was due to the introduction of high cyclomatic complexity
% in a method used to verify the password. In sum, \emph{Cross-Site Scripting} (CWE-79) should take part of the
% developers' preoccupations along with the other patterns that explicitly hinder
% maintainability because despite their low complexity these are patterns that
% still produce a considerable negative impact on the maintainability. Real
% efforts have been made in creating and incorporating coding standards and best
% practices in software production. However, security is still far from perfection
% and fixing vulnerabilities might hinder software maintainability. Thus, it is
% not only important to provide tool support to developers to help them apply these
% patterns without harming software maintainability, but also to create and redesign
% programming languages to facilitate developers improving software maintainability.
% We measured the effect sizes between both distribution -- before and after
% patches --- per language and security patterns using the Common Language Effect
% Size~\cite{cliff:1993} method. Effect sizes are of a low-medium level.
% Thus, considering the $0$ median and the low effect sizes, it is reasonable
% to affirm that there is a significant effect but it might only be
% recognized after a thoughtful data analysis.
% \subsubsection*{\textit{\textbf{RQ3} \textbf{What is the impact of security patches versus regular changes
% on the maintainability of open-source software?}}}
% \textit{Patching security vulnerabilities has a larger negative and positive impact
% than when fixing regular commits.}
% We found that $41.76\%$ of the security patches have a negative impact on the
% maintainability of open-source software (cf. Figure~\ref{fig:secvsreg}). While $21.04\%$ of regular commits have a negative impact on the maintainability. This raises
% a new trade-off when developers need to patch vulnerabilities on their projects
% because they might be affecting the software maintainability or even introducing
% new vulnerabilities. This study exhibits evidence that developers may have to
% reduce maintainability for the sake of security. We understand that developers
% should be able to produce secure code without affecting the maintainability of
% their projects. But there are still some concerns that need to be addressed:
% \begin{itemize}
% 	\item \textbf{Documentation} of software must provide developers with the
% 	best practices to implement secure and maintainable software. Unmaintainable
% 	code is difficult to test, analyze and re-use.
% 	\item\textbf{Programming Languages} should provide new design patterns to
% 	easily implement security patterns without endangering maintainability.
% 	Ultimately, new programming languages, both secure and maintainable by design,
% 	may be designed to help developers be highly productive when writing and
% 	maintaining  secure applications. One example is the need for designing new
% 	authorization mechanisms since these are one of the most complex security
% 	features to implement.
% 	\item \textbf{Software Developers} lack the knowledge to create secure and
% 	maintainable code. Mainly because academic curricula are not yet prepared
% 	to properly educate them. Universities should do their part on making
% 	this possible. Online services such as BCH have also a great potential to help
